Abstract

### 2025-01-23 output/Letter_to_editor_reviewers.Rmd
Version responses to Reviewer 1

We think the motivation and psychometric utility of our approach are more
related with the interpretation of the item parameters than with the latent
abilities.
To our understanding, the better interpretation of the item parameters that
results from our approach provides indeed a better insight into how the results
are used.
<!-- TODO: Complete -->

@@

@@

@@

@@

@@

@@

>>>>

We think the motivation and psychometric utility of our approach are more related with the interpretation of the item parameters than with the latent abilities.
To our understanding, the better interpretation of the item parameters that results from our approach provides indeed a better insight into how the results are used.
A multidimensional representation of the items provides benchmarks for comparing different latent trait levels, as discussed by @reckase_difficulty_1985.
Therefore, the more accurate such representation is, the better the latent trait interpretation will be.
<!-- TODO: Complete -->

@@

Our introduction is a bit less than four pages long.
We have provided concise paragraphs about the Multidimensional IRT model and its parameters, their assumptions, and our rationale to relax them.
Finally, a last paragraph presents the objectives of our research.
We honestly believe we have made a thorough effort to summarize the motivation of our research, and we find it difficult to shorten it further without losing relevant information to justify our work.
Therefore, we are sorry to say that we cannot comply with the reviewer's request to shorten the introduction.

@@

Our work is completely theoretical and thus uses a purely deductive method.
Therefore, we do not have any hypothesis to test, but the results of the formal derivations that we present, adressing the objective crearly described at the end of our introduction.

@@

The datasets we use in our paper are examples of item parameter estimates, taken from the literature.
As they only serve exemplification purposes, and are not the main focus of our article, we deem it more appropriate to describe their details in the corresponding section of the manuscript.
We have added a short reference to them in the introduction, to make sure their pertinence as examples is clear.

@@

We think the context makes it clear that we are referring to "within-item" multidimensionality.
In fact, it is clear from the manuscript that the results are trivial in the "between-item" multidimensionality case,
as the `r MDISC_SYM` parameter reduces to the only non-null component of the $`r DISCR_VECTOR`$ parameter.
However, we acknowledge this terminology is prevalent in the current MIRT literature;
therefore, we have now explained this in our "Model formulation" section.
The end of the second paragraph reads now as follows:

    > Third, as responses are not necessarily interpreted as correct or incorrect,
    parameter $`r INTERCEPT_PARAM`$ is said to be related to the item _location_
    instead of the item difficulty.
    Finally, it is worth noting that
    this formulation assumes *within-item multidimensionality* in general,
    which will be present whenever more than one component of $`r DISCR_VECTOR`$
    is non-null.
    However, the model can also account for *between-item multidimensionality*,
    if all items are unidimensional
    (i.e., their $`r DISCR_VECTOR`$ parameters only have one non-null component).

@@

Our last two paragraphs mention the most important limitations of our work.
We clearly demarcate the limited scope of our conclusions, and warn against overgeneralizing our results to other MIRT models or the common factor model.
We have put a great effort into writing our limitations in a way that is elegant and, at the same time, avoids cliché sentences that add little value.
Therefore, we would rather not add a separate section for limitations.

@@

Similarly to the previous comment, we believe that our conclusions are thoroughly and carefully elaborated in our "Discussion" section, and we would prefer not to add an explicit "Conclusions" header.

### 2025-01-23 output/Main_text.Rmd
Version responses to Reviewer 1

The main objective of this paper is thus to provide a generalization of
the MIRT parameters in two aspects: on one hand, in an oblique coordinate system
that can be soundly applied to the correlated trait case; on the other, to
items that yield monotonous response probabilities, but in the broader sense of
unchanging monotony. Additionally, we aim to provide a method for
graphically representing these items in the fashion of @ackerman_creating_1994
but taking the non-orthogonality of the coordinate axes into account.
To pursue these objectives, we first define the generalization of
the `r MODEL_ACRONYM` model and describe its scope of application,
which will require introducing some linear algebra concepts and results.
Then, we follow the method outlined by @reckase_difficulty_1985 and
@reckase_discriminating_1991 to obtain
the formulation of the multidimensional parameters.
We will introduce three different versions,
showing that either the original or one of the new ones is obtained
as a function of how we define and interpret the latent space.
Next, we discuss their properties,
propose a graphical representation method based on the new formulation,
and present two examples of application of our results.
We round up with a discussion about
the applicability and generalization of our results.

@@

We must stop here and make a few remarks. First, as we are not constraining
ourselves to cognitive traits,
we speak of _positive_ instead of _correct response_. This may imply giving
a correct response indeed, if we are talking about a cognitive test.
But it may mean endorsing an item (or a certain response option)
in a non-cognitive domain.
Second, although not explicitly stated, $`r DISCR_VECTOR`$ is a vector of
non-negative real numbers in @reckase_difficulty_1985
---a decrease in the probability of a correct response
with an increase in ability makes little sense].
In our case, $`r DISCR_VECTOR`$ is a vector of real numbers; the value
of the $`r DIM_INDEX`$-th component $`r DISCR_PARAM`$ can be negative if
the item is negatively-keyed for latent dimension $`r DIM_INDEX`$.
Finally, as responses are not necessarily interpreted as correct or incorrect,
parameter $`r INTERCEPT_PARAM`$ is said to be related to the item _location_
instead of the item difficulty.

>>>>

The main objective of this paper is thus to provide a generalization of
the MIRT parameters in two aspects: on one hand, in an oblique coordinate system
that can be soundly applied to the correlated trait case; on the other, to
items that yield monotonous response probabilities, but in the broader sense of
unchanging monotony. Additionally, we aim to provide a method for
graphically representing these items in the fashion of @ackerman_creating_1994
but taking the non-orthogonality of the coordinate axes into account.
To pursue these objectives, we first define the generalization of
the `r MODEL_ACRONYM` model and describe its scope of application,
which will require introducing some linear algebra concepts and results.
Then, we follow the method outlined by @reckase_difficulty_1985 and
@reckase_discriminating_1991 to obtain
the formulation of the multidimensional parameters.
We will introduce three different versions,
showing that either the original or one of the new ones is obtained
as a function of how we define and interpret the latent space.
Next, we discuss their properties,
propose a graphical representation method based on the new formulation,
and present two examples of application of our results.
These examples, from the cognitive and the non-cognitive domain respectively,
are taken from the MIRT literature:
One of them is a classic dataset, often used to exemplify
the use of the `r MODEL_ACRONYM` model [@reckase_multidimensional_2009];
the other one, an application in marketing research [@tezza_modelo_2018].
We round up with a discussion about
the applicability and generalization of our results.

@@

We must stop here and make a few remarks. First, as we are not constraining
ourselves to cognitive traits,
we speak of _positive_ instead of _correct response_. This may imply giving
a correct response indeed, if we are talking about a cognitive test.
But it may mean endorsing an item (or a certain response option)
in a non-cognitive domain.
Second, although not explicitly stated, $`r DISCR_VECTOR`$ is a vector of
non-negative real numbers in @reckase_difficulty_1985
---a decrease in the probability of a correct response
with an increase in ability makes little sense].
In our case, $`r DISCR_VECTOR`$ is a vector of real numbers; the value
of the $`r DIM_INDEX`$-th component $`r DISCR_PARAM`$ can be negative if
the item is negatively-keyed for latent dimension $`r DIM_INDEX`$.
Third, as responses are not necessarily interpreted as correct or incorrect,
parameter $`r INTERCEPT_PARAM`$ is said to be related to the item _location_
instead of the item difficulty.
Finally, it is worth highlighting that
this model generally assumes *within-item multidimensionality*,
which will be present whenever more than one component of $`r DISCR_VECTOR`$
is non-null.
However, the model can also account for *between-item multidimensionality*,
if the $`r DISCR_VECTOR`$ parameter only has one non-null component.
